Class differences in social networks: Evidence from a referral experiment

### 2025-07-30 writing/manuscript.tex
v1.1.2
Chapter 3 to-do: streamline language, rewrite introduction and conclusion

\section{Design} \label{s3}
We designed an online experiment to assess peer referral selection from an SES perspective and to evaluate the causal effect of providing a bonus to referral candidates. The experimental design consisted of two incentivized tasks administered in the following sequence to all participants: First, participants completed belief elicitation tasks about their own performance on the national university entry exam. Second, they completed the main referral task, nominating peers based on exam performance in two academic areas. Finally, participants reported beliefs about their referrals' performance and provided demographic information. This structure allowed us to measure both the accuracy of participants' beliefs and their referral behavior under controlled incentive conditions. Figure \ref{timeline} shows the experimental timeline, and detailed instructions are provided in Appendix \ref{instructions}.

\subsection{Performance measures}
To establish an objective basis for referral performance, we use national university entry exam scores (SABER 11). All Colombian high school students take the SABER 11 exam at the end of their final year as a requirement for university admission. These scores provide pre-existing, comparable measures of performance. By using existing administrative data, we also ensure that all eligible students have comparable performance measures.

The SABER 11 exam consists of five areas (critical reading, mathematics, natural sciences, social sciences, and English). We focus on critical reading and mathematics as these represent fundamental skills most relevant across academic programs and future employment. Critical reading evaluates competencies necessary to understand, interpret, and evaluate texts found in everyday life and broad academic fields (e.g., history), while mathematics assesses students' competency in using undergraduate level mathematical tools (e.g., reasoning in proportions, financial literacy). These together capture performance in comprehending and critically evaluating written material as well as reasoning and problem-solving abilities.

\subsection{Referral task}
The main experimental task involves making referrals among peers. For both exam areas (critical reading and mathematics), participants refer one peer they believe excels in that area. We provide an example question from the relevant exam area to clarify the skills that are being assessed. Participants type the name of their preferred candidate to make a referral. To avoid issues with recall, the interface provides autocomplete name and program suggestions from the administrative database (see Figure \ref{interface}).

We incentivize referrals using a piece rate payment structure. Referrers earn increasing payments as the percentile ranking of their recommendation increases (see Figure \ref{piece_rate_by_score}). We multiply the piece rate coefficient associated with the percentile rank by the actual exam scores of the recommendation to calculate earnings. This payment structure provides strong incentives to refer highly ranked peers with potential earnings going up to \$60 per referral. \footnote{Note that due to the selection into the university, the actual exam score distribution has limited variance. Below a certain threshold students cannot qualify for the institution and choose a lower ranked university, and above a certain threshold they have better options to choose from.}

\subsection{Bonus Treatment}
To examine how different incentive structures affect referral selection, we randomly assign a fixed bonus payment for students who get a referral. In the \textbf{Baseline} treatment, only the referrer can earn money based on their referral's performance. The \textbf{Bonus} treatment adds an additional fixed payment of \$25 to the peer who gets the referral. This payment is independent of the referred candidate's actual performance (see Table \ref{tab:treatments}).

\section{Sample, Incentives, and Procedure} \label{s4}
We invited all 4,417 undergraduate students who had completed their first year at the university at the time of recruitment to participate in our experiment. A total of 837 students participated in the data collection (19\% response rate). Our final sample consists of 734 individuals who referred peers with whom they had taken at least one class together, resulting in an 88\% success rate for the sample. We randomly allocated participants to either \textbf{Baseline} or \textbf{Bonus} treatments.

\subsection{Network characteristics}
We begin by describing the key features of the enrollment network for all participants. This network connects every participant in our sample with another university student if they have taken at least one course together at the time of data collection. By doing so, we construct the entire referral choice set for participants. We include in this dataset both the participant's and their potential candidate's individual characteristics, as well as the number of common courses they have taken together. Figure \ref{connections_class} describes the evolution of the enrollment network across the average number of network connections and the number of common courses taken with network members as participants progress through semesters.

\begin{figure}[H]
  \centering
  \caption{Network size and courses taken together by time spent at the university}\label{connections_class}
  \includegraphics[width=0.6\linewidth]{../figures/connections.png}
  \begin{tablenotes}
    \footnotesize
    \item[] \textit{Note:} This figure displays the average number of connections in blue and the average number of classes they have taken together with their connections in grey across semesters spent. The data shows an increase in the number of classes taken together as students progress in their programs, with the connections peaking around 7 semesters and dropping as certain students finish their bachelor's.
  \end{tablenotes}
\end{figure}

Having established the overall network structure, we now examine differences across SES groups. Are enrollment networks different across SES groups? We look at how the number of connections (network size) and number of courses taken together (tie strength) change across SES groups in Figure \ref{connections_by_ses}. Low- and middle-SES students have larger networks but take fewer courses together with network members, while high-SES students have smaller, denser networks. Specifically, both low- and middle-SES students have significantly larger networks than high-SES students ($t = 3.03, p = .003$ and $t = 2.49, p = .013$, respectively), but high-SES students take significantly more courses with their network members than both low- ($t = -3.70, p < .001$) and middle-SES ($t = -4.20, p < .001$).

\subsection{SES diversity in networks}\label{s:ses-diversity}
What are the diversity-related consequences of SES-driven differences across networks? In terms of network compositions, SES groups may connect with other SES groups at different rates than would occur randomly (Figure \ref{availability_lses}).\footnote{Because we estimate the share of SES groups in every individual network, we get very precise estimates of the actual means. However, it is important to note that these are not independent observations for each network. Estimates are precise because each network is a draw with replacement from the same pool of university population, from which we calculate the proportion of SES groups per individual network, and take the average over an SES group. Pooling over SES groups who are connected with similar others systematically reduces variance (similar to resampling in bootstrapping). For this reason we choose not reporting test results in certain sections including this one and focus on describing the relationships between SES groups.} Our results reveal modest deviations from university-wide SES composition across groups. Low-SES students have networks with 38.4\% low-SES peers compared to the university average of 34.3\%, middle-SES students connect with 52.9\% middle-SES peers versus the university average of 50.5\%, and high-SES students show 20.4\% high-SES connections compared to the university average of 15.3\%.

\begin{figure}[H]
  \centering
  \caption{Network shares of SES groups}\label{availability_lses}
  \includegraphics[width=0.6\linewidth]{../figures/availability_lses.png}
  \begin{tablenotes}
    \footnotesize
    \item[] \textit{Note:} This figure compares the network shares of SES groups in the networks of low-, middle-, and high-SES participants. Horizontal lines plot the university-wide shares of each SES group (Low: 35\%, Mid: 51\%, High: 14\%). While the share of low-SES peers in the network decreases as the SES of the group increases, the share of high-SES peers in the network increases.
  \end{tablenotes}
\end{figure}

At the same time, we observe much larger differences between SES groups in their connection patterns with other groups. Low-SES students connect with other low-SES students at higher rates than middle-SES students (38.4\% vs 31.4\%) and high-SES students (38.4\% vs 25.1\%). Conversely, high-SES students connect more with other high-SES students than both low-SES students (20.4\% vs 12.6\%) and middle-SES students (20.4\% vs 15.8\%). Middle-SES students are in between the two extreme patterns, connecting with middle-SES peers at higher rates than low-SES students (52.9\% vs 49.0\%) but lower rates than high-SES students (52.9\% vs 54.5\%). These findings indicate SES-based network segregation, with same-SES homophily patterns across groups.

Having examined basic network composition, we now turn to connection intensity. So far we have looked at the entire network without considering the intensity of connections between students. In our network dataset, this variable amounts to the number of classes taken together with peers. As we will see in the next section, referrals go to peers with whom participants have taken an average of 14 courses, implying the intensity of the connection matters. We begin by dissecting what the intensity means in our context. As students take more courses together, the proportion of peers from the same academic program quickly goes beyond 95\% (see Figure \ref{share_same_program_by_tie}). Similarly, the average network size drops very quickly from above 210 to below 50 (see Figure \ref{conns_by_tie}). Both results indicate that actual referral considerations originate from a much smaller pool of individuals from the same academic program.

\subsection{Program selection and SES diversity}
To understand the mechanisms driving these patterns, we examine program selection. Academic programs at this university use cost-based pricing, and typically less than 5\% of students receive any kind of scholarship \citep{diaz_peer_2025}. Based on this, we first calculate how much every program at the university is expected to cost students per year (see Figure \ref{fees_with_boxplot}). Considering that net minimum monthly wage stands at \$200 and the average Colombian salary around \$350, the cost differences between programs are large enough to make an impact on program selection. Is it the case that SES groups select into programs with financial considerations?

\subsection{Characterizing referrals}

We observe 1,342 referrals from our 734 participants in our final dataset. More than 90\% of these consist of participants referring for both areas of the national entry exam (see Appendix Table \ref{tab:referral_distribution}). While participants made one referral for Math and Reading parts of the exam, about 70\% of these referrals went to two separate individuals. We compare the outcomes across areas for unique referrals in Appendix Table \ref{tab:referral_by_area} and all referrals in Appendix Table \ref{tab:referral_area}. In both cases, we find no meaningful differences between referrals made for Math or Reading areas of the entry exam. As referrals in both exam areas come from the same referrer network, we pool referrals per participant and report their averages in our main analysis to avoid inflating statistical power in our comparisons.

First, referrals go to peers with whom the referrer has taken around 14 courses with on average, compared to almost 4 on average with others in their network (see Figure \ref{tie_hist}). This difference of 10.1 courses is significant ($t = 34.98$, $p < 0.001$), indicating that referrers choose individuals with whom they have stronger ties. While the median referral recipient has taken 12 courses together with the referrer, the median network member has shared only 2.8 courses. The interquartile range for referrals spans from 7.5 to 19.5 courses, compared to just 2.1 to 4.0 courses for the broader network, highlighting the concentration of referrals among peers with high social proximity and within same program (93\%).

\begin{figure}[H]
  \centering
  \caption{Courses taken together with network members and referrals}\label{tie_hist}
  \includegraphics[width=0.6\linewidth]{../figures/tie_hist.png}
  \begin{tablenotes}
    \footnotesize
    \item[] \textit{Note:} This figure compares the distributions of the number of courses taken together between referrers and their network members (orange) versus referrers and their chosen referral recipients (dark blue) for all 734 participants. 75\% of referral recipients having taken more than 7.5 courses together with the referrer, compared to only 25\% of network members. The distributions are significantly different (Kolmogorov-Smirnov test $D = 33.37$, $p < 0.001$).
  \end{tablenotes}
\end{figure}

Do referred individuals have different outcomes across treatments? We compare the performance, number of courses taken together, and SES shares of referred individuals between the \textbf{Baseline} and \textbf{Bonus} treatments in Table \ref{tab:referral_by_treatment}. While performance of referrals across Reading, Math, and GPA are similar across treatments, middle- and high-SES shares have significant differences. We find that referrals under the \textbf{Bonus} condition referred a higher proportion of high-SES individuals (13.5\% vs 8.8\%, $p=0.041$) and a lower proportion of middle-SES individuals on average (47.0\% vs 53.7\%, $p=0.072$). However, these differences do not appear to stem from systematic behavioral changes by any particular SES group of referrers, and the overall patterns remain largely consistent across treatments. The similarities in academic performance and number of courses taken together suggest that the core selection criteria—academic merit and social proximity—remain unchanged between conditions. For this reason, in the remainder of the paper, we report pooled results combining the averages of referral outcomes across treatments.

\subsection{Referral SES composition}
We first examine the overall SES compositions in referral selection. Referrals to low-SES peers constitute 37.9\% of all referrals, compared to 33.7\% low-SES representation in individual networks (see Figure \ref{fig:all_referral_rates}). This represents a modest over-representation of 4.3 percentage points. For middle-SES students, referrals constitute 51.0\% versus 51.4\% network representation, showing virtually no difference (-0.5 pp.). High-SES referrals account for 11.1\% compared to 14.9\% network share, an under-representation of 3.8 percentage points. While these patterns suggest some deviation from proportional representation—with slight over-referral to low-SES peers and under-referral to high-SES peers—the magnitudes are relatively modest. Overall, referral compositions are largely balanced and closely mirror the underlying network structure, with the largest deviation being less than 5 percentage points for any SES group.

We now shed more light on the referral behavior after having characterized how referrals were made. Particularly interesting is that referrals go to peers with whom the median participant took 12 courses, with an average of 14. By restricting the networks for courses taken above the median, we can get a snapshot of how the referral choice set actually looked for participants before making referral decisions. As discussed in Section \ref{s:ses-diversity}, taking more courses with network members increases the share of same-SES individuals for both low- and high-SES students, and we had explored program selection as a potential mechanism. In Figure \ref{t15availability}, we show the effects of network segregation on \textit{ex post} referral choice sets for each SES group. Network compositions above the median number of courses taken reveal strong segregation effects: Low-SES networks contain 44.5\% low-SES peers, higher than the 35\% university-wide share by 9.5 percentage points. Conversely, high-SES students are under-represented in low-SES networks at only 8.6\% average share, compared to the 14\% population share ($-5.4$ pp.). At the other extreme, high-SES networks show the reverse pattern with average low-SES share dropping to just 15.7\%, a 19.3 percentage point decrease relative to the university average. High-SES students have a same-SES concentration at 26.5\%, doubling their 14\% population share ($+12.5$ pp.). Middle-SES networks remain relatively balanced and closely track population proportions across all SES groups. Taken together, these suggest observed referral rates of SES groups may follow the network compositions above median number of courses taken together. We will test this formally by setting up a choice model where we can take into account individual differences in network compositions across SES, and try to identify SES biases that go beyond SES groups' availability in the choice sets.

To formally test for SES bias beyond network composition, we employ a choice modeling approach. We model a single referral outcome from mutually exclusive candidates, where our dependent variable outcome is multinomial distributed. Our design leverages the enrollment network to generate a dataset which includes alternative-specific variables for each referral decision, i.e., SES, courses taken together with the participant making the referral, as well as entry exam scores for not just the chosen alternative but all referral candidates. Using a conditional logit model on these data, we can identify whether an SES group has an aggregate bias controlling for each individual's unique enrollment network composition.

@@

@@

\begin{table}[H]
\centering
\begin{threeparttable}
\caption{SES bias in referral decisions by referrer SES group}\label{tab:ses-heterogeneity}
\begin{tabular*}{0.85\textwidth}{@{\extracolsep{\fill}}l c c c@{}}
\toprule
& \multicolumn{3}{c}{Referrer SES} \\
& Low & Middle & High \\
& (1) & (2) & (3) \\
\midrule
Low-SES candidate & 0.453*** & -0.019 & -0.710** \\
& (0.109) & (0.098) & (0.333) \\
High-SES candidate & -0.584*** & -0.255* & 0.001 \\
& (0.211) & (0.145) & (0.261) \\
\midrule
$\chi^2$ & 33.47 & 3.18 & 4.94 \\
Observations & 110,142 & 127,088 & 19,767 \\
Individuals & 301 & 366 & 67 \\
\bottomrule
\end{tabular*}
\begin{tablenotes}[flushleft]
\footnotesize
\item[] \textit{Note:} Individual-level clustered standard errors in parentheses. * $p < 0.1$, ** $p < 0.05$, *** $p < 0.01$. Each column represents a separate conditional logit regression estimated on the subsample of referrers from the indicated SES group. Coefficients represent log-odds of referring candidates from the specified SES group relative to referring middle-SES candidates. All models include individual fixed effects that control for each referrer's choice set composition.
\end{tablenotes}
\end{threeparttable}
\end{table}

Next, we include social proximity controls in our analysis. We proceed by adding the standardized number of courses taken together as a control in our specification and describe the results in Table \ref{tab:ses-heterogeneity-extended}. A one standard deviation increase in the number of courses taken together proves to be highly significant across all models, with coefficients ranging from 0.856 to 1.049, indicating that stronger social connections substantially increase the probability of referral. The high $\chi^2$ statistics suggest that these models explain considerably more variance than specifications without this control, highlighting the importance of courses taken together in referral decisions. Nevertheless, low-SES participants still show a strong same-SES bias relative to referring middle-SES peers at the average number of courses taken together. This same-SES bias is not observed among middle-SES or high-SES referrers, who also display no statistically significant bias toward low-SES candidates. No referrer group shows a positive bias for high-SES candidates relative to middle-SES candidates.

\begin{table}[H]
\centering
\begin{threeparttable}
\caption{SES bias in referral decisions by referrer SES group}\label{tab:ses-heterogeneity-extended}
\begin{tabular*}{0.85\textwidth}{@{\extracolsep{\fill}}l c c c@{}}
\toprule
& \multicolumn{3}{c}{Referrer SES} \\
& Low & Middle & High \\
& (1) & (2) & (3) \\
\midrule
Low-SES candidate & 0.348*** & -0.064 & -0.489 \\
& (0.123) & (0.115) & (0.337) \\
High-SES candidate & -0.366 & -0.165 & -0.140 \\
& (0.223) & (0.157) & (0.286) \\
Courses taken (z-score) & 0.856*** & 0.931*** & 1.049*** \\
& (0.035) & (0.037) & (0.126) \\
\midrule
$\chi^2$ & 626.15 & 636.10 & 71.43 \\
Observations & 110,142 & 127,088 & 19,767 \\
Individuals & 301 & 366 & 67 \\
\bottomrule
\end{tabular*}
\begin{tablenotes}[flushleft]
\footnotesize
\item[] \textit{Note:} Individual-level clustered standard errors in parentheses. * $p < 0.1$, ** $p < 0.05$, *** $p < 0.01$. Each column represents a separate conditional logit regression estimated on the subsample of referrers from the indicated SES group. Coefficients represent log-odds of referring candidates from the specified SES group relative to referring middle-SES candidates. All models include individual fixed effects that control for each referrer's choice set composition.
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

The evidence of a bias becoming significant when controlling for entry exam scores has a nuanced interpretation. While at the university-level, low-SES typically score lower in the entry exam, low-SES students appearing in high-SES networks are positively selected, scoring about 0.14 standard deviations higher than middle-SES students (see Appendix Table \ref{tab:network_connections}). Controlling for performance thus removes this positive selection and reveals the ``pure'' SES bias that was previously underestimated by above average performance of low-SES. Vice versa, high-SES in low-SES networks perform 0.12 standard deviations better than middle-SES students. The same bias was underestimated as high-SES candidates' better performance relative to middle-SES in the same networks provided a meritocratic justification for getting more referrals. Controlling for exam scores reveal that both high- and low-SES referrers have negative SES bias towards one another that operates independently of - and counter to - performance-based considerations. What makes interpretation difficult is that while biased against low-SES, high-SES referrers do not under any specification display a positive bias towards their in-group. For this final reason, we do not dig any further in this direction.

To conclude, we conduct joint significance tests, testing whether low- and high-SES regression coefficients are jointly different from middle-SES for each regression specification. For low-SES referrers, the joint test remains highly significant across all three specifications ($\chi^2 = 10.20$, $p = 0.006$ in the final model), indicating persistent SES bias across all specifications. In contrast, middle-SES referrers display no significant joint SES bias in any specification, with the test becoming increasingly non-significant as controls are added ($\chi^2 = 4.13$, $p = 0.127$ in the final model). High-SES referrers similarly show no significant joint SES bias across all three models ($\chi^2 = 4.28$, $p = 0.118$ in the final model). These results suggest that SES bias in referrals is primarily driven by low-SES. There is no sufficient evidence to conclude that middle- and high-SES referrers systematically discriminate against other-SES peers once we take into account the large differences in their network compositions due to program selection.

\begin{table}[H]
\centering
\begin{threeparttable}
\caption{SES bias in referral decisions by referrer SES group with academic performance controls}\label{tab:ses-heterogeneity-scores}
\begin{tabular*}{0.85\textwidth}{@{\extracolsep{\fill}}l c c c@{}}
\toprule
& \multicolumn{3}{c}{Referrer SES} \\
& Low & Middle & High \\
& (1) & (2) & (3) \\
\midrule
Low-SES candidate & 0.242** & -0.159 & -0.600* \\
& (0.123) & (0.114) & (0.327) \\
High-SES candidate& -0.445** & -0.274* & -0.345 \\
& (0.222) & (0.157) & (0.287) \\
Courses taken (z-score) & 0.859*** & 0.948*** & 1.043*** \\
& (0.036) & (0.038) & (0.118) \\
Entry exam (candidate z-score) & 0.607*** & 0.587*** & 0.883*** \\
& (0.052) & (0.047) & (0.111) \\
\midrule
$\chi^2$ & 789.87 & 756.06 & 120.54 \\
Observations & 110,142 & 127,088 & 19,767 \\
Individuals & 301 & 366 & 67 \\
\bottomrule
\end{tabular*}
\begin{tablenotes}[flushleft]
\footnotesize
\item[] \textit{Note:} Individual-level clustered standard errors in parentheses. * $p < 0.1$, ** $p < 0.05$, *** $p < 0.01$. Each column represents a separate conditional logit regression estimated on the subsample of referrers from the indicated SES group. Coefficients represent log-odds of referring candidates from the specified SES group relative to referring middle-SES candidates. All models include individual fixed effects that control for each referrer's choice set composition.
\end{tablenotes}
\end{threeparttable}
\end{table}

>>>>

We construct enrollment networks using administrative data to map social connections at the university.  For each participant, we identify all other undergraduate students with whom they have taken at least one course and create their individual network of university connections. The size of this network depends on how many different students a participant has encountered through coursework, while the intensity of connection is measured by the number of courses taken together. This approach provides a complete picture of each participant's social environment at the university, including detailed characteristics (i.e., SES, academic program, performance) for both the participant and every person in their network.

\textcolor{cyan}{why don't we introduce number of programs and student count here as background?}

\section{Empirical Specification} \label{s3}

To formally test for SES bias, we use a choice modeling approach. We observe a single referral outcome from mutually exclusive candidates. Our design leverages the enrollment network to generate a dataset which includes alternative-specific variables for each referral decision, i.e., SES, courses taken together with the participant making the referral, as well as entry exam scores for not just the chosen alternative but all referral candidates. Using a conditional logit model, we can identify whether an SES group has an aggregate bias controlling for each individual's unique enrollment network composition.

@@

\section{Design} \label{s4}
We designed an online experiment to assess peer referral selection from an SES perspective and to evaluate the causal effect of providing a bonus to referral candidates. The experimental design consisted of two incentivized tasks administered in the following sequence: First, participants completed belief elicitation tasks about their own performance on the national university entry exam. Second, they completed the main referral task, nominating peers based on exam performance in two academic areas. Finally, participants reported beliefs about their referrals' performance and provided demographic information. This structure allowed us to measure both the accuracy of participants' beliefs and their referral behavior under controlled incentive conditions. Figure \ref{timeline} shows the experimental timeline, and detailed instructions are provided in Appendix \ref{instructions}.

\subsection{Performance measures}
To establish an objective basis for referral performance, we use national university entry exam scores (SABER 11). All Colombian high school students take the SABER 11 exam at the end of their final year as a requirement for university admission. The scores from this exam provide pre-existing, comparable measures of performance. By using existing administrative data, we also ensure that all eligible students have comparable performance measures.

The exam consists of five areas (critical reading, mathematics, natural sciences, social sciences, and English). We focus on critical reading and mathematics as these represent two independent and fundemental skills. Critical reading evaluates competencies necessary to understand, interpret, and evaluate texts found in everyday life and broad academic fields (e.g., history). Mathematics assesses students' competency in using undergraduate level mathematical tools (e.g., reasoning in proportions, financial literacy). These together capture performance in comprehending and critically evaluating written material as well as reasoning and problem-solving abilities.

\subsection{Referral task}
The main task involves making referrals among peers. For both exam areas (critical reading and mathematics), participants refer one peer they believe excels in that area. We provide an example question from the relevant exam area to clarify the skills that are being assessed. Participants type the name of their preferred candidate to make a referral. To avoid issues with recall, the interface provides autocomplete name and program suggestions from the administrative database (see Figure \ref{interface}).

We incentivize referrals using a piece rate payment structure. Referrers earn increasing payments as the percentile ranking of their referral increases (see Figure \ref{piece_rate_by_score}). We multiply the piece rate coefficient associated with the percentile rank by the actual exam scores of the referral to calculate earnings. This payment structure provides strong incentives to refer highly ranked peers with potential earnings going up to \$60 per referral. \footnote{Note that due to the selection into the university, the actual exam score distribution has limited variance. Below a certain threshold students cannot qualify for the institution and choose a lower ranked university, and above a certain threshold they have better options to choose from.}

\subsection{Bonus Treatment}
To examine how different incentive structures affect referral selection, we randomly assign a fixed bonus payment for students who get a referral. In the \textbf{Baseline} treatment, only the participants can earn money based on their referral's performance. The \textbf{Bonus} treatment adds an additional fixed payment of \$25 to the peer who gets the referral. This payment is independent of the referral's actual performance (see Table \ref{tab:treatments}).

\section{Sample, Incentives, and Procedure} \label{s5}
We invited all 4,417 undergraduate students who had completed their first year at the university at the time of recruitment to participate in our experiment. A total of 837 students participated in the data collection (19\% response rate). Our final sample consists of 734 individuals who referred peers with whom they had taken at least one class together (88\% success rate). We randomly allocated participants to either \textbf{Baseline} or \textbf{Bonus} treatments.

\subsection{Network characteristics}
We begin by describing the key features of the enrollment networks. On average, participants connect with 175 other students, and take an average of 3.62 courses together. Figure \ref{connections_class} shows how network characteristics vary by students' time at the university: both the number of connections (network size) and the number of courses taken together (connection intensity) change as participants progress through their studies.

\begin{figure}[H]
  \centering
  \caption{Network size and courses taken together by time spent at the university}\label{connections_class}
  \includegraphics[width=0.6\linewidth]{../figures/connections.png}
  \begin{tablenotes}
    \footnotesize
    \item[] \textit{Note:} 
    This figure displays the average number of connections in blue and the average number of courses taken together with connections in grey across semesters completed. Network size (nb. of connections) peaks around 7 semesters before declining as students graduate. Connection intensity (nb. of courses taken) has an increasing trend.
  \end{tablenotes}
\end{figure}

We now examine how connection intensity relates to network size and composition. First, if two students take more courses together, it is very likely that they are in the same academic program. We plot this relationship in Figure \ref{share_same_program_by_tie}: As students take more than 5 courses together, the share of students in their enrollment network from the same academic program quickly exceeds 90\%. Second, because students sort into specialized academic programs, increases in courses taken together should result in decreases in connections. We plot this relationship in Figure \ref{conns_by_tie}: As students take more than 5 courses together, the size of their enrollment network drops dramatically from above 210 to below 50. These patterns reveal that while participants' overall networks are large with relatively few courses taken together on average, they spend most of their time at the university within smaller, more intensive groups of peers from the same academic program.

\subsection{Referral characteristics}

Participants made one referral for math and one referral for the reading part of the university entry exam from their enrollment networks. We observe 1,342 referrals from 734 participants in our final dataset. More than 90\% of these consist of participants referring for both exam areas (see Appendix Table \ref{tab:referral_distribution}). About 70\% of these referrals go to two separate individuals. We compare the outcomes across exam areas for referrals only going to separate individuals in Appendix Table \ref{tab:referral_by_area} and all referrals in Appendix Table \ref{tab:referral_area}. In both cases, we find no meaningful differences between referrals made for Math or Reading areas of the entry exam. As referrals in both exam areas come from the same enrollment network, we pool referrals per participant and report their averages in our main analysis to avoid inflating statistical power in our comparisons.

First, referrals go to peers with whom the referrer has taken around 14 courses with on average, compared to almost 4 on average with others in their network (see Figure \ref{tie_hist}). This difference of 10.1 courses is significant ($t = 34.98$, $p < 0.001$), indicating that referrers choose individuals with whom they have stronger ties. While the median referral recipient has taken 12 courses together with the referrer, the median network member has shared only 2.8 courses. The interquartile range for referrals spans from 7.5 to 19.5 courses, compared to just 2.1 to 4.0 courses for the broader network, highlighting the concentration of referrals among peers with high social proximity. In addition, 93\% of referrals go to students within same program.

\begin{figure}[H]
  \centering
  \caption{Courses taken together with network members and referrals}\label{tie_hist}
  \includegraphics[width=0.6\linewidth]{../figures/tie_hist.png}
  \begin{tablenotes}
    \footnotesize
    \item[] \textit{Note:} This figure compares the distributions of the number of courses taken together between referrers and their network members (orange) versus referrers and their chosen referral recipients (dark blue) for all 734 participants. 75\% of referral recipients take more than 7.5 courses together with the referrer, compared to only 25\% of network members. The distributions are significantly different (Kolmogorov-Smirnov test $D = 33.37$, $p < 0.001$).
  \end{tablenotes}
\end{figure}

Do referred individuals have different outcomes across treatments? We compare the performance, number of courses taken together, and SES shares of referred individuals between the \textbf{Baseline} and \textbf{Bonus} treatments in Table \ref{tab:referral_by_treatment}. While performance of referrals across Reading, Math, and GPA are similar across treatments, middle- and high-SES shares have significant differences. We find that referrals under the \textbf{Bonus} condition referred a higher proportion of high-SES individuals (13.5\% vs 8.8\%, $p=0.041$) and a lower proportion of middle-SES individuals on average (47.0\% vs 53.7\%, $p=0.072$). The similarities in academic performance and number of courses taken together suggest that performance and contact intensity drive referrals regardless of treatment. For this reason, in the remainder of the paper, we report pooled results combining the averages of referral outcomes across treatments.

@@

\begin{table}[H]
\centering
\begin{threeparttable}
\caption{SES bias in referral decisions by referrer SES group}\label{tab:ses-heterogeneity}
\begin{tabular*}{0.85\textwidth}{@{\extracolsep{\fill}}l c c c@{}}
\toprule
& \multicolumn{3}{c}{Referrer SES} \\
& Low & Middle & High \\
& (1) & (2) & (3) \\
\midrule
Low-SES referral & 0.453*** & -0.019 & -0.710** \\
& (0.109) & (0.098) & (0.333) \\
High-SES referral & -0.584*** & -0.255* & 0.001 \\
& (0.211) & (0.145) & (0.261) \\
\midrule
$\chi^2$ & 33.47 & 3.18 & 4.94 \\
Observations & 110,142 & 127,088 & 19,767 \\
Individuals & 301 & 366 & 67 \\
\bottomrule
\end{tabular*}
\begin{tablenotes}[flushleft]
\footnotesize
\item[] \textit{Note:} Individual-level clustered standard errors in parentheses. * $p < 0.1$, ** $p < 0.05$, *** $p < 0.01$. Each column represents a separate conditional logit regression estimated on the subsample of referrers from the indicated SES group. Coefficients represent log-odds of referring from the specified SES group relative to referring middle-SES candidates. All models include individual fixed effects that control for each referrer's choice set composition.
\end{tablenotes}
\end{threeparttable}
\end{table}

Next, we include social proximity controls in our analysis. We proceed by adding the standardized number of courses taken together as a control in our specification and describe the results in Table \ref{tab:ses-heterogeneity-extended}. A one standard deviation increase in the number of courses taken together proves to be highly significant across all models, with coefficients ranging from 0.856 to 1.049, indicating that intensity of contact substantially increase the probability of referral. The high $\chi^2$ statistics suggest that these models explain considerably more variance than specifications without this control, highlighting the predictive power of courses taken together in referral decisions. Nevertheless, low-SES participants still show a strong same-SES bias relative to referring middle-SES peers at the average number of courses taken together. This same-SES bias is not observed among middle-SES or high-SES referrers, who also display no statistically significant bias toward low-SES candidates. No referrer group shows a positive bias for high-SES candidates relative to middle-SES candidates.

\begin{table}[H]
\centering
\begin{threeparttable}
\caption{SES bias in referral decisions by referrer SES group}\label{tab:ses-heterogeneity-extended}
\begin{tabular*}{0.85\textwidth}{@{\extracolsep{\fill}}l c c c@{}}
\toprule
& \multicolumn{3}{c}{Referrer SES} \\
& Low & Middle & High \\
& (1) & (2) & (3) \\
\midrule
Low-SES referral & 0.348*** & -0.064 & -0.489 \\
& (0.123) & (0.115) & (0.337) \\
High-SES referral & -0.366 & -0.165 & -0.140 \\
& (0.223) & (0.157) & (0.286) \\
Courses taken (z-score) & 0.856*** & 0.931*** & 1.049*** \\
& (0.035) & (0.037) & (0.126) \\
\midrule
$\chi^2$ & 626.15 & 636.10 & 71.43 \\
Observations & 110,142 & 127,088 & 19,767 \\
Individuals & 301 & 366 & 67 \\
\bottomrule
\end{tabular*}
\begin{tablenotes}[flushleft]
\footnotesize
\item[] \textit{Note:} Individual-level clustered standard errors in parentheses. * $p < 0.1$, ** $p < 0.05$, *** $p < 0.01$. Each column represents a separate conditional logit regression estimated on the subsample of referrers from the indicated SES group. Coefficients represent log-odds of referring  from the specified SES group relative to referring middle-SES candidates. All models include individual fixed effects that control for each referrer's choice set composition.
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

The evidence of a bias becoming significant when controlling for entry exam scores has a nuanced interpretation. While at the university-level, low-SES typically score lower in the entry exam, low-SES students appearing in high-SES networks are positively selected, scoring about 0.14 standard deviations higher than middle-SES students (see Appendix Table \ref{tab:network_connections}). Controlling for performance thus removes this positive selection and reveals the ``pure'' SES bias that was previously underestimated by above average performance of low-SES. Vice versa, high-SES in low-SES networks perform 0.12 standard deviations better than middle-SES students. The same bias was underestimated as high-SES candidates' better performance relative to middle-SES in the same networks provided a meritocratic justification for getting more referrals. Controlling for exam scores reveal that both high- and low-SES referrers have negative SES bias towards one another that operates independently of - and counter to - performance-based considerations. What makes interpretation difficult is that while biased against low-SES, high-SES referrers do not (under any specification) display a positive bias towards their in-group. For this final reason, we do not dig any further in this direction.

To conclude, we conduct joint significance tests, testing whether low- and high-SES regression coefficients are jointly different from middle-SES for each regression specification. For low-SES referrers, the joint test remains highly significant across all three specifications ($\chi^2 = 10.20$, $p = 0.006$ in the final model), indicating persistent SES bias across all specifications. In contrast, middle-SES referrers display no significant joint SES bias in any specification, with the test becoming increasingly non-significant as controls are added ($\chi^2 = 4.13$, $p = 0.127$ in the final model). High-SES referrers similarly show no significant joint SES bias across all three models ($\chi^2 = 4.28$, $p = 0.118$ in the final model). These results suggest that SES bias in referrals is primarily driven by low-SES. There is no sufficient evidence to conclude that middle- and high-SES referrers systematically discriminate against other-SES peers. Naturally, this null result occurs once we take into account the potential differences in the network compositions of each SES group. In the next section, we explore the differences in the network compositions of different SES groups.
